---
title: 'Exam: Introduction to probability and statistics'
author: "M2:Masco 2025-2026"
date: "28-11-2025"
output: pdf_document
---

## Exercise 1: 

In a game of chance where a coin is flipped, you win when you get heads and you lose when you get tails.
A coin is randomly chosen from a bag containing two types of biased coins, A and B.
The probability of winning with a type-A coin is $80\%$, and the probability of winning with a type-B coin is $10\%$.
We also know that the probability of choosing a type-A coin is $30\%$ and that of choosing a type-B coin is $70\%$.

\begin{enumerate}
    \item A coin is chosen at random and flipped once. What is the probability of losing?
    \item Given that you win, what is the probability that the chosen coin was of type A?
\end{enumerate}


## Exercise 2: 
We model the number of \emph{spikes} during an experiment by a random variable $X$ following 
a Poisson distribution with unknown parameter $\Theta=\theta>0$. We write:
\[
P(X = x \mid \Theta=\theta) = e^{-\theta}\,\frac{\theta^{x}}{x!}, \qquad x = 0,1,2,\ldots
\]

Based on prior knowledge of the phenomenon, we know that the random variable $\Theta$ is 
discrete and takes the values $\{0.1,\,0.5,\,1,\,2\}$ with associated probabilities 
$\{0.25,\,0.25,\,0.4,\,0.1\}$.

\begin{enumerate}
    \item In a first experiment, we observe $2$ spikes. Give the posterior distribution of $\Theta$.
    \item Determine the posterior mean of $\Theta$.
    \item Compute the posterior probability that $\Theta$ is smaller than $1$.
    \item The experiment is repeated and we now observe $1$ spike. Assuming the observations are independent,
    determine the posterior distribution of $\Theta$ based on both experiments.
\end{enumerate}


## Exercise 3:

Weather modification, or cloud seeding, consists of treating individual clouds or storm systems with various inorganic and organic materials in the hope of increasing precipitation. The dataset \texttt{clouds.csv} contains two variables giving the logarithm of precipitation measured in acre-foot (about $1233\,\text{m}^3$) for $26$ unseeded clouds (\texttt{unseeded}) and $26$ seeded clouds (\texttt{seeded}).

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "C:/Users/labo-maths/Downloads")
```


```{r}
clouds <- read.csv("clouds.csv")
X <- clouds$unseeded
Y <- clouds$seeded
head(clouds)
```


\begin{itemize}
\item[1.] Explain what each line of the above code does.
\end{itemize}

You want to examine whether cloud seeding has an impact on precipitation.

\begin{itemize}
  \item[2.] Propose a procedure to investigate this (state the null hypothesis, the alternative, and the assumptions required, etc.).
\end{itemize}

You decide to run the following tests


```{r}
#a. test 1
shapiro.test(X)

#b. test 2
shapiro.test(Y)

#c. test 3
var.test(X,Y)

#d. test 4
t.test(X,Y, var.equal = TRUE)
```
\begin{enumerate}
\item[3.] Explain what each test is for and why we choose \texttt{var.equal = TRUE} in test 4.
\item[4.] What are your conclusions?
\end{enumerate}

## Exercise 4 
\begin{enumerate}

\item Here are the observed values of two samples (considered independent):
\[
A = (176, 391, 314, 248, 373, 393, 330), \qquad
B = (307, 310, 374, 343, 360, 280).
\]
We perform a Wilcoxon rank sum test. What is the value of the test statistic? Show your calculations in detail.

\item If we add 200 to each element of $A$ and $B$, does the value of the test statistic change?

\item We observe a sample $(X_1,\ldots,X_n)$ of independent random variables such that $X_i \sim \mathrm{Binomial}(10,\theta)$, $\theta\in(0,1)$.  
We want to estimate $\theta$.  
The proposed estimator is
\[
\hat{\theta} = \frac{n}{10(n-2)}\,\bar{X} - \frac{2}{n},
\qquad
\bar{X} = \frac{1}{n}\sum_{i=1}^n X_i.
\]
Compute the bias and variance of this estimator. Comment on your result.

\item Compare and briefly discuss the Bayesian and frequentist approaches in statistical inference.

\item An experimenter claims: "To reject a null hypothesis of equal means, it is enough to collect sufficiently large samples." Comment on this claim.

\item Briefly explain what Bayesian ANOVA is and what it does.

\item What is the Bayes factor for two models, \(M_1\) and \(M_2\)? How do we interpret \(BF_{12} = 4\)?

\item If \(BF_{01} = 3\) and \(BF_{12} = 2\), which model is better: \(M_0\) or \(M_2\)?


\end{enumerate}



